---
title: "Publication-worthy Sleuth Analysis"
output: html_notebook
---

```{r}
library(ggplot2)
library(sleuth)
library(cowplot)
```

"""
Import the reference human genome list of transcripts to gene list
"""

```{r}
# Import the reference human genome
mart <- biomaRt::useMart(biomart = 'ENSEMBL_MART_ENSEMBL', dataset = 'hsapiens_gene_ensembl', host = 'ensembl.org')

# Set the reference object to index transcipt ID to the gene ID
t2g <- biomaRt::getBM(attributes = c("ensembl_transcript_id", "transcript_version", "ensembl_gene_id", "external_gene_name", "description", "transcript_biotype"), mart = mart)
t2g <- dplyr::rename(t2g, target_id = ensembl_transcript_id, ens_gene = ensembl_gene_id, ext_gene = external_gene_name)
t2g <- dplyr::select(t2g, c('target_id', 'ens_gene', 'ext_gene'))
```

"""
Import the list of all kallisto transcript count .csv file locations,
as denoted by the folder and sample_table.tsv file
"""

```{r}
# Open the folder containing the kallisto results
sample_id <- dir(file.path('~/Desktop', 'RNAseq', "kallisto"))
kal_dirs <- sapply(sample_id, function(id) file.path('~/Desktop/RNAseq', "kallisto",id))
```

```{r}
# Associate each kallisto output with its associated conditions
s2c <- read.table(file.path('~/Desktop/RNAseq', "sample_table.tsv"), header = TRUE, stringsAsFactors=FALSE)
s2c <- dplyr::rename(s2c, sample = Run_s)
s2c <- dplyr::mutate(s2c, path = kal_dirs)
s2c
```

"""


```{r}
# Condition 1, uninfected RD or A549 cells at either 33C or 37C
RD_A549_Mock <- s2c[s2c$strain=="Mock", ][c("sample","cell", "temperature", "path")]

# Condition 2, RD cells Mock or infected with WT at either 33C or 37C
RD_WT_or_Mock <- s2c[s2c$cell=="RD" & s2c$strain!="I278V", ][c("sample","strain", "temperature", "path")]

# Condition 3, RD cells with either WT or I278V at 33C or 37C
RD_WT_or_I278V <- s2c[s2c$cell=="RD" & s2c$strain!="Mock", ][c("sample","strain", "temperature", "path")]

# Condition 4, A549 cells Mock or infected with WT at either 33C or 37C
A549_WT_or_Mock <- s2c[s2c$cell=="A549" & s2c$strain!="I278V", ][c("sample","strain", "temperature", "path")]

# Condition 5, A549 cells with either WT or I278V at 33C or 37C
A549_WT_or_I278V <- s2c[s2c$cell=="A549" & s2c$strain!="Mock", ][c("sample","strain", "temperature", "path")]

# Condition 6, RD cells, 33C or 37C
RD_Mock <- s2c[s2c$strain=="Mock" & s2c$cell=="RD", ][c("sample", "cell", "temperature", "path")]

# Condition 7, RD and A549 cells, 37C
RD_A549_37C_Mock <- s2c[s2c$strain=="Mock" & s2c$temperature=="37", ][c("sample","cell", "temperature", "path")]

# Condition 8, RD and A549 cells, 37C
RD_WT <- s2c[s2c$strain=="WT" & s2c$cell=="RD", ][c("sample","cell", "temperature", "path")]

# Condition 9, RD and A549 at 33C
RD_A549_33C_Mock <- s2c[s2c$strain=="Mock" & s2c$temperature=="33",][c("sample","cell","temperature", "path")]
```

"""
Create the sleuth objects from each set of conditions
"""

```{r}
# Condition 1
so_RD_A549_Mock <- sleuth_prep(RD_A549_Mock, target_mapping = t2g, extra_bootstrap_summary = TRUE)
# Condition 2
so_RD_WT_or_Mock <- sleuth_prep(RD_WT_or_Mock, target_mapping = t2g, extra_bootstrap_summary = TRUE)
# Condition 3
so_RD_WT_or_I278V <- sleuth_prep(RD_WT_or_I278V, target_mapping = t2g, extra_bootstrap_summary = TRUE)
# Condition 4
so_A549_WT_or_Mock <- sleuth_prep(A549_WT_or_Mock, target_mapping = t2g, extra_bootstrap_summary = TRUE)
# Condition 5
so_A549_WT_or_I278V <- sleuth_prep(A549_WT_or_I278V, target_mapping = t2g, extra_bootstrap_summary = TRUE)
# Condition 6
so_RD_Mock <- sleuth_prep(RD_Mock, target_mapping = t2g, extra_bootstrap_summary = TRUE, aggregation_column = 'ens_gene')
# Condition 7 
so_RD_A549_37C_Mock <- sleuth_prep(RD_A549_37C_Mock, target_mapping = t2g, extra_bootstrap_summary = TRUE)
# Condition 8 
so_RD_WT_Mock <- sleuth_prep(RD_WT, target_mapping = t2g, extra_bootstrap_summary = TRUE)
# Condition 9 
so_RD_mixed <- sleuth_prep(RD_mixed_condition, target_mapping = t2g, extra_bootstrap_summary = TRUE)
# Condition 10
```





"""
Fit the conditions to either the full model or only individual conditions
"""

```{r}
RD_WT_or_Mock$strain
```


```{r}
# Fit the sleuth object of condition 1 to both parameters or only cell or temperature
so_RD_A549_Mock <- sleuth_fit(so_RD_A549_Mock, ~temperature + cell, 'full')
so_RD_A549_Mock <- sleuth_fit(so_RD_A549_Mock, ~temperature, 'reduced_cell')
so_RD_A549_Mock <- sleuth_fit(so_RD_A549_Mock, ~cell, 'reduced_temperature')
```

```{r}
# Fit the sleuth object of condition 2 to both parameters or only strain or temperature
so_RD_WT_or_Mock <- sleuth_fit(so_RD_WT_or_Mock, ~strain + temperature, 'full')
so_RD_WT_or_Mock <- sleuth_fit(so_RD_WT_or_Mock, ~temperature, 'reduced_strain')
so_RD_WT_or_Mock <- sleuth_fit(so_RD_WT_or_Mock, ~strain, 'reduced_temperature')
```

```{r}
# Fit the sleuth object of condition 3 to both parameters or only strain or temperature
so_RD_WT_or_I278V <- sleuth_fit(so_RD_WT_or_I278V, ~temperature + strain, 'full')
so_RD_WT_or_I278V <- sleuth_fit(so_RD_WT_or_I278V, ~temperature, 'reduced_strain')
so_RD_WT_or_I278V <- sleuth_fit(so_RD_WT_or_I278V, ~strain, 'reduced_temperature')
```

```{r}
# Fit the sleuth object of condition 4 to both parameters or only strain or temperature
so_A549_WT_or_Mock <- sleuth_fit(so_A549_WT_or_Mock, ~temperature + strain, 'full')
so_A549_WT_or_Mock <- sleuth_fit(so_A549_WT_or_Mock, ~temperature, 'reduced_strain')
so_A549_WT_or_Mock <- sleuth_fit(so_A549_WT_or_Mock, ~strain, 'reduced_temperature')
```

```{r}
# Fit the sleuth object of condition 5 to both parameters or only strain or temperature
so_A549_WT_or_I278V <- sleuth_fit(so_A549_WT_or_I278V, ~temperature + strain, 'full')
so_A549_WT_or_I278V <- sleuth_fit(so_A549_WT_or_I278V, ~temperature, 'reduced_strain')
so_A549_WT_or_I278V <- sleuth_fit(so_A549_WT_or_I278V, ~strain, 'reduced_temperature')
```

```{r}
# Fit the sleuth object of condition 6 to both parameters or only strain or temperature
so_RD_Mock <- sleuth_fit(so_RD_Mock, ~temperature, 'full')
```

```{r}
# Fit the sleuth object of condition 7 to the parameter of cell type
so_RD_A549_37C_Mock <- sleuth_fit(so_RD_A549_37C_Mock, ~cell, 'full')
# Fit the sleuth object of condition 8 to the parameter of temperature
so_RD_WT_Mock <- sleuth_fit(so_RD_WT_Mock, ~cell, 'full')
```

```{r}
# Fit the sleuth object of condition 9 to the parameter of A549 cells vs RD at 37
so_RD_mixed <- sleuth_fit(so_RD_mixed, ~cell, 'full')
```

"""
Perform likelihood ratio test on the five sleuth conditions
"""

```{r}
# Perform lrt on condition 1 with either temperature or cell type
so_RD_A549_Mock <- sleuth_lrt(so_RD_A549_Mock, 'reduced_cell', 'full')
so_RD_A549_Mock <- sleuth_lrt(so_RD_A549_Mock, 'reduced_temperature', 'full')
```

```{r}
# Perform lrt on condition 2 with either temperature or cell type
so_RD_WT_or_Mock <- sleuth_lrt(so_RD_WT_or_Mock, 'reduced_strain', 'full')
so_RD_WT_or_Mock <- sleuth_lrt(so_RD_WT_or_Mock, 'reduced_temperature', 'full')
```

```{r}
# Perform lrt on condition 3 with either temperature or cell type
so_RD_WT_or_I278V <- sleuth_lrt(so_RD_WT_or_I278V, 'reduced_strain', 'full')
so_RD_WT_or_I278V <- sleuth_lrt(so_RD_WT_or_I278V, 'reduced_temperature', 'full')
```

```{r}
# Perform lrt on condition 4 with either temperature or cell type
so_A549_WT_or_Mock <- sleuth_lrt(so_A549_WT_or_Mock, 'reduced_strain', 'full')
so_A549_WT_or_Mock <- sleuth_lrt(so_A549_WT_or_Mock, 'reduced_temperature', 'full')
```

```{r}
# Perform lrt on condition 5 with either temperature or cell type
so_A549_WT_or_I278V <- sleuth_lrt(so_A549_WT_or_I278V, 'reduced_strain', 'full')
so_A549_WT_or_I278V <- sleuth_lrt(so_A549_WT_or_I278V, 'reduced_temperature', 'full')
```

"""
Perform the Wald test on the 5 sleuth condition-dependent objects
for either of the parameters
"""

```{r}
# Perform the Wald test on condition 1 for cell type and temperature
so_RD_A549_Mock <- sleuth_wt(so_RD_A549_Mock, "cellRD")
so_RD_A549_Mock <- sleuth_wt(so_RD_A549_Mock, "temperature")
```

```{r}
# Perform the Wald test on condition 2 for strain type and temperature
so_RD_WT_or_Mock <- sleuth_wt(so_RD_WT_or_Mock, "strainWT")
so_RD_WT_or_Mock <- sleuth_wt(so_RD_WT_or_Mock, "temperature")
```

```{r}
# Perform the Wald test on condition 3 for strain type and temperature
so_RD_WT_or_I278V <- sleuth_wt(so_RD_WT_or_I278V, "strainWT")
so_RD_WT_or_I278V <- sleuth_wt(so_RD_WT_or_I278V, "temperature")
```

```{r}
# Perform the Wald test on condition 4 for strain type and temperature
so_A549_WT_or_Mock <- sleuth_wt(so_A549_WT_or_Mock, "strainWT")
so_A549_WT_or_Mock <- sleuth_wt(so_A549_WT_or_Mock, "temperature")
```

```{r}
# Perform the Wald test on condition 5 for strain type and temperature
so_A549_WT_or_I278V <- sleuth_wt(so_A549_WT_or_I278V, "strainWT")
so_A549_WT_or_I278V <- sleuth_wt(so_A549_WT_or_I278V, "temperature")
```

```{r}
# Perform the Wald test on condition 6 for just temperature
so_RD_Mock <- sleuth_wt(so_RD_Mock, "temperature")
```

```{r}
# Perform the Wald test on condition 7 for just cell type
so_RD_A549_37C_Mock <- sleuth_wt(so_RD_A549_37C_Mock, "cellRD")
```

```{r}
# Perform the Wald test on condition 8 for just cell type
so_RD_WT_Mock <- sleuth_wt(so_RD_WT_Mock, 'temperature')
```

```{r}
# Perform the Wald test on condition 9 for RD at 33 or all A549
so_RD_mixed <- sleuth_wt(so_RD_mixed, 'cellRD')
```

"""
Save the sleuth objects for downstream analysis
"""


```{r}
# Save the sleuth object for both Wald analysis on condition 1
sleuth_table_gene <- sleuth_results(so_RD_A549_Mock, 'temperature', 'wt', show_all = FALSE)
write.csv(sleuth_table_gene, "~/Documents/EVD68_adaptation/data/RNAseq/Sleuth_files/RD_A549_Mock_temperature.csv")

sleuth_table_gene <- sleuth_results(so_RD_A549_Mock, 'cellRD', 'wt', show_all = FALSE)
write.csv(sleuth_table_gene, "~/Documents/EVD68_adaptation/data/RNAseq/Sleuth_files/RD_A549_Mock_cell.csv")
```

```{r}
# Save the sleuth object for both Wald analysis on condition 2
sleuth_table_gene <- sleuth_results(so_RD_WT_or_Mock, 'temperature', 'wt', show_all = FALSE)
write.csv(sleuth_table_gene, "~/Desktop/RNAseq/RD_WT_or_Mock_temperature.csv")

sleuth_table_gene <- sleuth_results(so_RD_WT_or_Mock, 'strainWT', 'wt', show_all = FALSE)
write.csv(sleuth_table_gene, "~/Desktop/RNAseq/RD_WT_or_Mock_strain.csv")
```

```{r}
# Save the sleuth object for both Wald analysis on condition 3
sleuth_table_gene <- sleuth_results(so_RD_WT_or_I278V, 'temperature', 'wt', which_model = 'full', show_all = TRUE)
write.csv(sleuth_table_gene, "~/Desktop/RNAseq/RD_WT_or_I278V_temperature.csv")

sleuth_table_gene <- sleuth_results(so_RD_WT_or_I278V, 'strainWT', 'wt', show_all = TRUE)
write.csv(sleuth_table_gene, "~/Desktop/RNAseq/RD_WT_or_I278V_strain.csv")
```

```{r}
# Save the sleuth object for both Wald analysis on condition 4
sleuth_table_gene <- sleuth_results(so_A549_WT_or_Mock, 'temperature', 'wt', show_all = FALSE)
write.csv(sleuth_table_gene, "~/Documents/EVD68_adaptation/data/RNAseq/Sleuth_files/so_A549_WT_or_Mock_temperature.csv")

sleuth_table_gene <- sleuth_results(so_A549_WT_or_Mock, 'strainWT', 'wt', show_all = FALSE)
write.csv(sleuth_table_gene, "~/Documents/EVD68_adaptation/data/RNAseq/Sleuth_files/so_A549_WT_or_Mock_strain.csv")
```

```{r}
# Save the sleuth object for both Wald analysis on condition 5
sleuth_table_gene <- sleuth_results(so_A549_WT_or_I278V, 'temperature', 'wt', show_all = FALSE)
write.csv(sleuth_table_gene, "~/Documents/EVD68_adaptation/data/RNAseq/Sleuth_files/A549_WT_or_I278V_temperature.csv")

sleuth_table_gene <- sleuth_results(so_A549_WT_or_I278V, 'strainWT', 'wt', show_all = FALSE)
write.csv(sleuth_table_gene, "~/Documents/EVD68_adaptation/data/RNAseq/Sleuth_files/A549_WT_or_I278V_strain.csv")
```

```{r}
# Save the sleuth object for both Wald analysis on condition 6
sleuth_table_gene <- sleuth_results(so_RD_Mock, 'temperature', 'wt', show_all = TRUE, pval_aggregate = T)
write.csv(sleuth_table_gene, "~/Documents/EVD68_adaptation/data/RNAseq/Sleuth_files/RD_Mock_temperature_agg.csv")
```

```{r}
# Save the sleuth object for both Wald analysis on condition 7
sleuth_table_gene <- sleuth_results(so_RD_A549_37C_Mock, 'cellRD', 'wt', show_all = FALSE)
write.csv(sleuth_table_gene, "~/Documents/EVD68_adaptation/data/RNAseq/Sleuth_files/RD_A549_37C_Mock_cell.csv")
```

```{r}
# Save the sleuth object for both Wald analysis on condition 8
sleuth_table_gene <- sleuth_results(so_RD_WT_Mock, 'temperature', 'wt', show_all = FALSE)
write.csv(sleuth_table_gene, "~/Documents/EVD68_adaptation/data/RNAseq/Sleuth_files/RD_WT_temperature.csv")
```

```{r}
# Save the sleuth object for the Wald analysis on condition 9
sleuth_table_gene <- sleuth_results(so_RD_mixed, 'cellRD', 'wt', show_all=FALSE)
write.csv(sleuth_table_gene, "~/Documents/EVD68_adaptation/data/RNAseq/Sleuth_files/RD_A549_mixed.csv")
```

"""
Visualize the sleuth objects
"""

```{r}
sleuth_live(so_RD_Mock)
```


